Understanding the impact of electronically-delivered parenting advice

### 2019-03-09 paper/cogsci2019/cogsci2019.tex
update participants info and e2 results

Young children spend a large portion of their waking time at play,
variously manipulating objects, exploring their environment, and
interacting with caregivers and peers. Playing with objects allows them
to discover hidden object properties and relations, and to build a
causal understanding of how objects interact (e.g., Schulz \& Bonawitz,
2007). Meanwhile, play also gives children an opportunity to set and
achieve goals (e.g., build a tower) and to practice a wide range of
motor skills (e.g., stacking) that will help them navigate the world
(Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help children
learn about human relationships, both through imitation of adult
behaviors and by experiencing and learning to process emotional events
such as failures (Singer et al., 2006). Of course, young children are
rarely playing in isolation: caregivers often provide encouragement and
guidance while scaffolding a child's play (Kaye, 1970; Wood, Bruner, \&
Ross, 1976). Early parenting practices have been shown to play an
important role in shaping the future outcomes of young children (Hart \&
Risley, 1995; Heckman, 2006). While many interventions have trouble
reaching the parents of very young children, digital delivery of
parenting advice by way of

@@

With the widespread use of smartphones and tablets worldwide,
digitally-delivered interventions could address many of the logistical
barriers that have limited scaling up face-to-face delivery methods. A
review of 11 studies verified that digital delivery is a promising means
of effectively distributing interventions (Breitenstein, Gross, \&
Christophersen, 2014). However, the parent and child outcomes assessed
in the review (e.g.infant positive behaviors, satisfaction, emotional
symptoms etc.) did not address the nature or quality of parent-infant
interactions at a detailed level, like if the interventions lead to
children paying more attention, or vocabulary changes in parents'
language usages. Although digitally-delivered ac-tivities are designed
to promote learning and cognitive development, it is unclear how they
might affect these dimensions of parent-child interactions. Thus, we
want to conduct this experiment to explore if and how digital
scaffolding of activities affect the social and linguistic
characteristics of parent-child ineractions. The quality of parent-child
interactions can be measured by both the social engagement of parents
(e.g., joint attention to objects in the environment)(Bigelow, MacLean,
\& Proctor, 2004) and the quality of language (e.g., vocabulary
diversity) (Malvern, Richards, Chipere, \& Durán, 2004).

@@

The video of each session was manually coded for episodes of joint
attention using the Datavyu software (Team, 2014). The video taken at
floor level was coded by default, but the other video was referred to if
the participants were not visible or if there was technical difficulty
with the first camera. Each session's video was coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. Parental bids for joint attention
were defined as any attempt to initiate joint attention (i.e labeling,
pointing, or otherwise drawing attention to an object) that did not
result in passive or coordinated joint attention. If more than 3 seconds
elapsed between bids, they were coded as separate attempts. An episode
of joint attention was considered passive if both participants visually
focused on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. Episodes
joint attention were considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent. This could be looks to the
parent's face, gestures, vocalizations, or turn-taking. If either
participant looked away for less than 3 seconds and then returned to the
same object, it was coded as part of the same period of joint attention.
A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders had a reliability of ICC = 0.80
with 95\% confident interval (CI) = {[}0.57,0.92{]} (p \textless{} 0.05)
for number of parent bids for JA; ICC = 0.20 with 95\% CI =
{[}-0.26,0.58{]} for number of passive JA episodes; ICC = 0.66 with 95\%
CI = {[}0.32,0.85{]} (p \textless{} 0.05) for number of coordinated JA
episodes; ICC = 0.24 with 95\% CI = {[}-0.21,0.61{]} for total duration
of passive JA episodes, and ICC = 0.62 with 95\% CI = {[}0.27,0.83{]} (p
\textless{} 0.05) for total duration of passive (coordinated??) JA
episodes.

@@

84 infants aged 12-24 months and their parents participated in the same
museum as Experiment 1. We included infants who were exposed to English
at least 75 percent of the time or who were exposed less but whose
participating parent reported that they primarily speak English with
their child at home. {[}specific demo info{]}

@@

The procedure for Experiment 2 matched that of Experiment 1, except that
parents in the Control Video condition watched a control video before
the play session. Consistent with the No-Video control condition in
Experiment 1, parents in the Control Video condition were told to play
with their child as they would at home, and were not given additional
instructions.

@@

@@

@@

>>>>

Young children spend a large portion of their waking time at play,
variously manipulating objects, exploring their environment, and
interacting with caregivers and peers. Playing with objects allows them
to discover hidden object properties and relations, and to build a
causal understanding of how objects interact (e.g., Schulz \& Bonawitz,
2007). Meanwhile, play also gives children an opportunity to set and
achieve goals (e.g., build a tower) and to practice a wide range of
motor skills (e.g., stacking) that will help them navigate the world
(Singer, Golinkoff, \& Hirsh-Pasek, 2006). Social play can help children
learn about human relationships, both through imitation of adult
behaviors and by experiencing and learning to process emotional events
such as failures (Singer et al., 2006). Of course, young children are
rarely playing in isolation: caregivers often provide encouragement and
guidance while scaffolding a child's play (Kaye, 1970; Wood, Bruner, \&
Ross, 1976). Indeed, parenting practices early in childhood have been
shown to play an important role in shaping the future outcomes of young
children (Hart \& Risley, 1995; Heckman, 2006). While interventions
often have trouble reaching many parents of very young children, the
proliferation of mobile devices offers a good avenut for the digital
delivery of parenting advice (Breitenstein, Gross, \& Christophersen,
2014).

@@

With the widespread use of smartphones and tablets worldwide,
digitally-delivered interventions could address many of the logistical
barriers that have limited scaling up face-to-face delivery methods. A
review of 11 studies verified that digital delivery is a promising means
of effectively distributing interventions (Breitenstein et al., 2014).
However, the parent and child outcomes assessed in the review
(e.g.infant positive behaviors, satisfaction, emotional symptoms etc.)
did not address the nature or quality of parent-infant interactions at a
detailed level, like if the interventions lead to children paying more
attention, or vocabulary changes in parents' language usages. Although
digitally-delivered ac-tivities are designed to promote learning and
cognitive development, it is unclear how they might affect these
dimensions of parent-child interactions. Thus, we want to conduct this
experiment to explore if and how digital scaffolding of activities
affect the social and linguistic characteristics of parent-child
ineractions. The quality of parent-child interactions can be measured by
both the social engagement of parents (e.g., joint attention to objects
in the environment)(Bigelow, MacLean, \& Proctor, 2004) and the quality
of language (e.g., vocabulary diversity) (Malvern, Richards, Chipere, \&
Durán, 2004).

@@

The video of each session was manually coded for episodes of joint
attention using the Datavyu software (Team, 2014). The video taken at
floor level was coded by default, but the other video was referred to if
the participants were not visible or if there was technical difficulty
with the first camera. Each session's video was coded for episodes of
coordinated joint attention, episodes of passive joint attention, and
parental bids for joint attention. Parental bids for joint attention
were defined as any attempt to initiate joint attention (i.e labeling,
pointing, or otherwise drawing attention to an object) that did not
result in passive or coordinated joint attention. If more than 3 seconds
elapsed between bids, they were coded as separate attempts. An episode
of joint attention was considered passive if both participants visually
focused on an object for a minimum of 3 seconds but the child did not
acknowledge the parent. If either participant looked away from the
object for less than 3 seconds and then returned to the same object it
was considered part of the same period of joint attention. Episodes
joint attention were considered coordinated if both participants
visually focused on an object for a minimum of 3 seconds and at some
point in the interaction the child indicated awareness of interaction
with some overt behavior toward the parent. This could be looks to the
parent's face, gestures, vocalizations, or turn-taking. If either
participant looked away for less than 3 seconds and then returned to the
same object, it was coded as part of the same period of joint attention.
A second coder independently coded a third of the videos (i.e., 20 of
the 60 videos, approximately equally distributed across ages) to
establish reliability. The two coders had a reliability of ICC = 0.80
with 95\% confident interval (CI) = {[}0.57,0.92{]} (p \textless{} 0.05)
for number of parent bids for JA; ICC = 0.20 with 95\% CI =
{[}-0.26,0.58{]} for number of passive JA episodes; ICC = 0.66 with 95\%
CI = {[}0.32,0.85{]} (p \textless{} 0.05) for number of coordinated JA
episodes; ICC = 0.24 with 95\% CI = {[}-0.21,0.61{]} for total duration
of passive JA episodes, and ICC = 0.62 with 95\% CI = {[}0.27,0.83{]} (p
\textless{} 0.05) for total duration of coordinated JA episodes.

@@

84 infants (F = 37, M = 47) aged 12-24 months (42 12-17.9 month-olds, 42
18-24 month-olds) and their parents participated in the same museum as
Experiment 1. We included infants who were exposed to English at least
75 percent of the time or who were exposed less but whose participating
parent reported that they primarily speak English with their child at
home. Forty nine\% of participants (n = 41) had been exposed to two or
more languages as indicated by their parent. Parents identified their
children as White (n = 39), Asian (n = 20), African American/Black (n =
1), Biracial (n = 9), other (n = 7), or declined to state (n = 8).
Sixteen parents reported their child was of Hispanic origin. Parents
tended to be highly- educated, with reports of highest level of
education ranging from completed high school (n = 0), some college (n =
5), four-year college (n = 28), some graduate school (n = 2), to
completed graduate school (n = 35) or declined to state (n = 14).

@@

The procedure for Experiment 2 matched that of Experiment 1, except that
parents in the Control Video condition watched a control video before
the play session. Consistent with the No-Video control condition in
Experiment 1, parents in the Control Video condition were told to play
with their child as they would at home, and were not given additional
instructions. The coding procedure also matched that of Experiment 1. A
second coder independently coded a third of the videos (i.e., 26 of the
84 videos, approximately equally distributed across ages) to establish
reliability. The two coders had a reliability of ICC = 0.80 with 95\%
confident interval (CI) = {[}0.60,0.90{]} (p \textless{} 0.05) for
number of parent bids for JA; ICC = 0.74 with 95\% CI = {[}0.59,0.87{]}
(p \textless{} 0.05) for number of passive JA episodes; ICC = 0.78 with
95\% CI = {[}0.58,0.90{]} (p \textless{} 0.05) for number of coordinated
JA episodes; ICC = 0.72 with 95\% CI = {[}0.46,0.86{]} (p \textless{}
0.05) for total duration of passive JA episodes, and ICC = 0.88 with
95\% CI = {[}0.75,0.94{]} (p \textless{} 0.05) for total duration of
coordinated JA episodes.

@@

Parents' child-directed speech was transcribed and processed in the same
way as in Experiment 1.

@@

We fit a mixed-effects linear regression predicting TTR and MTLD as a
function of age (scaled and 0-centered) and condition with a random
intercept per video using lme4 (Bates et al., 2015). There was
significantly lower TTR in the Video condition (mean: 0.38) than in the
Control condition (mean: 0.47, \(\beta=-.09\), t(8) = 3.16,
\emph{p}\textless{}.05). There was no significant effect of age. A
similar mixed-effects linear regression instead predicting MTLD found no
significant effects of age or condition. Figure 1 shows the mean of each
lexical diversity measure (TTR and MTLD) by condition.

We also conducted similar regressions predicting the number of word
tokens and types, finding no significant effects of of age or condition.

@@

{\centering \includegraphics{figs/e2ja-graphs-1}

\caption[Mean number of bids and episodes of Joint Attention in Experiment 2]{Mean number of bids and episodes of Joint Attention in Experiment 2.}\label{fig:e2ja-graphs}
\end{figure}
\end{CodeChunk}

{\centering \includegraphics{figs/e2ja-graphs-pass-coord-1}

\caption[Average number of passive and coordinated episodes of JA in Experiment 2]{Average number of passive and coordinated episodes of JA in Experiment 2.}\label{fig:e2ja-graphs-pass-coord}
\end{figure}
\end{CodeChunk}
